---
title: "6737932_SIRE506_Assignment02"
output:
  word_document: default
  html_document: default
date: "2024-09-08"
editor_options:
  markdown:
    wrap: 72
---

```{r}
library(readr)
train <- read_csv("Documents/SIRE/506/SIRE506_Assignment02/Dataset/train.csv")
```

Note: Package "readr" is installed to read CSV file by using read_csv()
function.

# Summary Statistic for Titanic by Survival Status

A comprehensive overview of major characteristic of 891
passengers during titanic diaster has been analyzed and categorized base
on survival status. Figure 1 reveal that their are 342 survivors whereas there are 549 non-survivors, the
number of female passengers show the significant survivals rate as 68% which is greater than male(32%).
The reflection of passengers whose traveling none or at least one
Siblings/Spouses aboard (SibSp) or Parents/Children aboard (Patch) had
the higher potential of survival rates. Moreover, passenger in
first-class got the highest survivals rate with 40% followed by 35% of
thired-class and least survival at second-class which is 25% of
passengers. The median of age has been calculated and indicates central
value at 28 years old, however, the unknown ages can affects the
analysis of age in the survival pattern. The fare paid of the survived
passenger tended to higher than who not survived. The presence of these
finding suggests that fare paid could be one of the factors that
determined the survival of passengers which could related to the passengers class
because the higher paid can indicate the greater service and location in the cabin, therefore, those factors can influence the chance of servive throughout this diaster.

Note: Essential R packages that has been install are "gt", "dplyr", and
"gtsummary".

```{r}
library(gt)
library(dplyr)
library(gtsummary)
summary_table <- train%>% select(Survived, Pclass, Sex, Age, SibSp, Parch, Fare) %>% tbl_summary(by = Survived)
summary_table_gt <- summary_table %>%
  as_gt() %>%
  tab_caption(caption = "Figure 1: Summary Statistic for Titanic by Survival Status") %>%
  tab_options(table.width = pct(100))
summary_table_gt <- summary_table_gt %>%
  tab_footnote(
    footnote = " Explanatory Notes:
                 Survived: Passenger did not survive (0), Survived passenger (1). 
                 Pclass: Passenger class (1, 2, or 3).
                 Sex: Gender of the passenger. 
                 Age: Age of the passenger.
                 SibSp: Number of siblings and spouses aboard the Titanic.
                 Parch: Number of children and parents aboard the Titanic.
                 Fare: The price paid (in pounds) for the ticket.",
    locations = cells_column_labels(columns = everything())
  )
summary_table_gt

```


## Identify outliers and the spread of continuoues variables

In order to visuailzing the distribution of continuous variables and
identifying outliers, the box plot of the age distribution of passengers
(figure 2.1) and the fare paid of passengers (figure 2.2) are use to
observed the correlation coefficient of demographic in titanic dataset.
Additionally, the box plot analysis are use to support and clarify the
data from figure 1.

```{r}
boxplot(Age ~ Survived, 
        data = train, 
        main = "Figure 2.1: Comparison of Age Distributions by Survival",
        xlab = "Survival Status",
        ylab = "Age",
        names = c("Not Survived", "Survived"),
        col = c("gray", "lightblue"))
boxplot(Fare ~ Survived, 
        data = train, 
        main = "Figure 2.2: Comparison of Fare Distributions by Survival",
        xlab = "Survival Status",
        ylab = "Fare Amount",
        names = c("Not Survived", "Survived"),
        col = c("gray", "lightblue"))
```
